Level-Set method on \texttt{p4est}

### 2014-12-04 paper/tex/scaling_results.tex
Started with reinitialization scaling

\subsection{Semi-Lagrangian}
To test the scalability of the Semi-Lagrangian algorithm \ref{alg:semi-lagrangian} we consider a slightly modified version of the Enright's rotation test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}, i.e. we advect a sphere of radius 0.35 located at $(0.4, 0.4, 0.4)$ with a divergent free velocity field given by:
\bean
u(x,y,z) &=& 2\sin(\pi x)^2 \sin(2\pi y) \sin(2 \pi z), \\
v(x,y,z) &=& -\sin(\pi y)^2 \sin(2\pi x) \sin(2 \pi z),\\
w(x,y,z) &=& -\sin(\pi z)^2 \sin(2\pi x) \sin(2 \pi y). 
\eean
To understand the effect of CFLnumber on the scalability of the algorithm we perform one step of the Semi-Lagrangian algorithm for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$. We also perform the test for two set of initial girds, a small grid with $l_\max = 10$ and a larger grid with $l_\max = 12$. After one advection step, each of these grids have approximately 15M and 255M nodes, respectively. 
\begin{table}[htbp]
	\begin{minipage}{.5\linewidth}
		\begin{center}
		\begin{tabular}{ccccccc}
			\hline
				CFL & 16 & 32 & 64 & 128 & 256 & 512 \\
				\hline
				1   & 2 & 2 & 3 & 3 & 3 & 3 \\
				10  & 3 & 3 & 3 & 3 & 3 & 3 \\
				100 & 6 & 6 & 6 & 6 & 6 & 6 \\
			\hline
		\end{tabular}
		\caption*{(a) $l_\max = 10$}
		\end{center}
	\end{minipage}%
	\begin{minipage}{.5\linewidth}
		\begin{center}
		\begin{tabular}{ccccccc}
			\hline
				CFL & 128 & 256 & 512 & 1024 & 2048 & 4096 \\
			\hline
				1   & 3 & 3 & 3 & 3 & 3 & 3 \\
				10  & 3 & 3 & 4 & 4 & 4 & 4 \\
				100 & 6 & 6 & 6 & 6 & 6 & 7 \\
			\hline
		\end{tabular}
		\caption*{(b) $l_\max = 12$}
		\end{center}
	\end{minipage}	
	\caption{Number of iterations required for grid construction in algorithm \ref{alg:semi-lagrangian} for the rotation test on a (a) level-10 and (b) level-12 octree with approximately 15M and 255M nodes, respectively. Note how the iteration count increases with the CFL number but is almost independent of number of processors. The slight dependence of iteration count on number of processors is most likely due to dependence of round-off errors on the number of processors. Nonetheless, close examination of generated octrees (data not shown) reveals that they are identical and independent of the number of processors used to perform the test.}
	\label{tab:semilagrangian}
\end{table}

@@

An increase in the CFL number has two effects on the algorithm. First, a larger fraction of departure points land in domains of remote processors. Moreover, these points are potentially dispersed across a larger number of processors. This means both the number MPI messages and their volume increases for larger CFL numbers. Second, as more points are shipped to remote processors for interpolation, there is a greater chance that interpolation load is unbalanced across processors. This is specially true for regions of space in which streamlines cluster. Both factors contribute to decrease the scalability of the algorithm at larger CFL numbers. To shed more light on this, we have recorded a complete history of communication pattern in the interpolation step.

\begin{figure}[htbp]
	\begin{center}
		\subfigure[Semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_1_SemiLagrangian.pdf}}
		\subfigure[Semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_10_SemiLagrangian.pdf}}
		\subfigure[Semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_100_SemiLagrangian.pdf}}
		\\
		\subfigure[Interpolation, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_1_Interpolation.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_10_Interpolation.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_100_Interpolation.pdf}}
	\end{center}
	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-10 octree with approximately 13M cells and 15M nodes. Top row: scaling of various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$. Bottom row: breakdown of various components of the interpolation phase for the same CFL numbers. Note that the maximum time has been scaled with the number of iterations required to build the tree (c.f. table \ref{tab:semilagrangian}).}
	\label{fig:semilagrangian_small}
\end{figure}

@@

% \subsection{Reinitialization} \label{section::scaling_reinitialization}

% The reinitialization procedure presented in section \ref{section::reinitialization} necessitates the second order derivatives of the level-set function for the second order discretization in space. The finite difference schemes we implemented rely on the description of the neighborhood of each vertex in the adaptive tree environment. This information can either be computed on the fly or stored in a buffer. Fetching the neighbors information is rather time consuming and was observed to improve the overall scalability of the algorithm by increasing the local computation load. However, it slowed down the procedure significantly and therefore we choose to present the buffered version of the implementation. If one was to be limited in memory, constructing the neighborhood information on the fly would be a desirable option and would lead to better scaling, but in our case memory is not a limiting factor.

% The scaling results are presented in figure \ref{fig::scaling_reinitialization} and demonstrate satisfying scalability. Taking for reference the timing with $128$ processes, we observe efficiencies of $0.877$ with $1024$ processes, $0.763$ with $2048$ processes and $0.609$ with $4096$ processes.

% \begin{figure}[ht!]
% \begin{center}
% \includegraphics[width=.7\textwidth]{figures/scaling_reinitialization_1st_time_2nd_space_with_buffer.pdf}
% \caption{Scalability of the reinitialization procedure presented in section \ref{section::reinitialization} and analyzed in section \ref{section::scaling_reinitialization}.} \label{fig::scaling_reinitialization}
% \end{center}
% \end{figure}

% \subsection{Poisson solver}

>>>>

\subsection{Semi-Lagrangian}
To test the scalability of the Semi-Lagrangian algorithm \ref{alg:semi-lagrangian} we consider a slightly modified version of the Enright's rotation test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}, i.e. we advect a sphere of radius 0.35 located at $(0.4, 0.4, 0.4)$ with a divergent free velocity field given by:
\bean
u(x,y,z) &=& 2\sin(\pi x)^2 \sin(2\pi y) \sin(2 \pi z), \\
v(x,y,z) &=& -\sin(\pi y)^2 \sin(2\pi x) \sin(2 \pi z),\\
w(x,y,z) &=& -\sin(\pi z)^2 \sin(2\pi x) \sin(2 \pi y). 
\eean
To understand the effect of CFLnumber on the scalability of the algorithm we perform one step of the Semi-Lagrangian algorithm for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$. We also perform the test for two set of initial girds, a small grid with $l_\text{max} = 10$ and a larger grid with $l_\text{max} = 12$. After one advection step, each of these grids have approximately 15M and 255M nodes, respectively. 
\begin{table}[htbp]
	\begin{minipage}{.5\linewidth}
		\begin{center}
		\begin{tabular}{|c|cccccc|}
			\hline
				CFL & 16 & 32 & 64 & 128 & 256 & 512 \\
				\hline
				1   & 2 & 2 & 3 & 3 & 3 & 3 \\
				10  & 3 & 3 & 3 & 3 & 3 & 3 \\
				100 & 6 & 6 & 6 & 6 & 6 & 6 \\
			\hline
		\end{tabular}
		\caption*{(a) $l_\text{max} = 10$}
		\end{center}
	\end{minipage}%
	\begin{minipage}{.5\linewidth}
		\begin{center}
		\begin{tabular}{|c|cccccc|}
			\hline
				CFL & 128 & 256 & 512 & 1024 & 2048 & 4096 \\
			\hline
				1   & 3 & 3 & 3 & 3 & 3 & 3 \\
				10  & 3 & 3 & 4 & 4 & 4 & 4 \\
				100 & 6 & 6 & 6 & 6 & 6 & 7 \\
			\hline
		\end{tabular}
		\caption*{(b) $l_\text{max} = 12$}
		\end{center}
	\end{minipage}	
	\caption{Number of iterations required for grid construction in algorithm \ref{alg:semi-lagrangian} for the rotation test on a (a) level-10 and (b) level-12 octree with approximately 15M and 255M nodes, respectively. Note how the iteration count increases with the CFL number but is almost independent of number of processors. The slight dependence of iteration count on number of processors is most likely due to dependence of round-off errors on the number of processors. Nonetheless, close examination of generated octrees (data not shown) reveals that they are identical and independent of the number of processors used to perform the test.}
	\label{tab:semilagrangian}
\end{table}

@@

An increase in the CFL number has two effects on the algorithm. First, a larger fraction of departure points land in domains of remote processors. Moreover, these points are potentially dispersed across a larger number of processors. This means both the number MPI messages and their volume increases for larger CFL numbers. Second, as more points are shipped to remote processors for interpolation, there is a greater chance that interpolation load is unbalanced across processors. This is specially true for regions of space in which streamlines cluster. Both factors contribute to decrease the scalability of the algorithm at larger CFL numbers. 
\begin{figure}[htbp]
	\begin{center}
		\subfigure[Semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_1_SemiLagrangian.pdf}}
		\subfigure[Semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_10_SemiLagrangian.pdf}}
		\subfigure[Semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_100_SemiLagrangian.pdf}}
		\\
		\subfigure[Interpolation, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_1_Interpolation.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_10_Interpolation.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_100_Interpolation.pdf}}
	\end{center}
	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-10 octree with approximately 13M cells and 15M nodes. Top row: scaling of various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$. Bottom row: breakdown of various components of the interpolation phase for the same CFL numbers. Note that the maximum time has been scaled with the number of iterations required to build the tree (c.f. table \ref{tab:semilagrangian}).}
	\label{fig:semilagrangian_small}
\end{figure}

@@

To better understand the importance of CFL number on the scalability, we have recorded a complete history of communication pattern in the interpolation step. Figure \ref{fig:communication_4096} illustrates the effects of the CFL number on three different metrics, namely number of interpolation points\footnote{Note that this includes both local points and those queried by other processors}, $N_p$, number of sent and received messages, $N_m = S + R$, and the total communication volume, $V_m$ in mega bytes (MB), for $p=4096$ processors. Moreover, these values are reported for the first (top row) and last (bottom row) iterations of the Semi-Lagrangian algorithm. There are several points to make. First, increasing the CFL number greatly increases the load imbalance as evident in the spread of data in figure \ref{fig:communication_4096_pf}. This is because at higher CFL numbers, it is more likely that some processors receive larger portions of backtracked points. Second, increasing the CFL number increases both the communication volume and its spread across processors (c.f. figure \ref{fig:communication_4096_vf}. Interestingly, however, the number of sent and received messages seems to remain unaffected by the CFL number. The bottom row of figure \ref{fig:communication_4096} illustrates better balance both in the computation and communication volume in the last iteration of the Semi-Lagrangian algorithm. This can be explained by noting that as the algorithm converges to the final grid, partitioning of the grid essentially becomes equivalent to partitioning of the backtracked points. Detailed information about the load balancing and communication pattern is listed in table \ref{tab:communication}.
\begin{figure}[htbp]
	\begin{center}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/point_first_4096.pdf}  \label{fig:communication_4096_pf}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/message_first_4096.pdf}\label{fig:communication_4096_mf}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/volume_first_4096.pdf} \label{fig:communication_4096_vf}}
		\\
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/point_last_4096.pdf}   \label{fig:communication_4096_pl}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/message_last_4096.pdf} \label{fig:communication_4096_ml}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/volume_last_4096.pdf}  \label{fig:communication_4096_vl}}		
	\end{center}
	\caption{Performance indicators of the first (top row) and last (bottom row) iterations of the Semi-Lagrangian algorithm for the advection on 4096 processors for $\text{CFL} = 1$ (\drawdiamond{black}), $\text{CFL} = 10$ (\drawsquare{red}), and $\text{CFL} = 100$ (\drawcircle{blue}). Increasing the CFL number greatly causes load imbalance in the amount of computation done during interpolation (a) and increases the communication volume (c). However the CFL number does not seem to appreciably affect the number of messages sent by processors (b). During the last Semi-Lagrangian iteration the initial grid, $G_0$ (c.f. algorithm \ref{alg:semi-lagrangian}), is very close to the final grid. As a result load imbalance is considerably improved (d). Curiously, however, communication pattern does not seem to be affected by much between first and last iterations (e-f).}
	\label{fig:communication_4096}
\end{figure}

\begin{table}[htbp]
	\centering
	\resizebox{\columnwidth}{!}{%
	\begin{tabular}{|c|c|cccc|cccc|}
	\hline
	\multirow{2}{*}{CFL} & \multirow{2}{*}{Metric} & \multicolumn{4}{|c|}{First Iteration} & \multicolumn{4}{|c|}{Last Iteration} \\
	\cline{3-10}
											 & 												 & min & max & avg & stddev 					   & min & max & avg & stddev 						\\
	\hline
	\multirow{3}{*}{1}   & $N_p$        & 6.67E+04 & 1.59E+05 & 7.57E+04 & 4.86E+03 & 6.69E+04 & 1.55E+05 & 7.57E+04 & 4.73E+03 \\
										   & $N_m$ 			  & 18 			 & 387 		  & 47.55 	 & 31.72 		& 18 			 & 392 			& 47.55 	 & 31.19 		\\
										   & $V_m$ (MB)   & 4.01E-01 & 2.93E+00 & 7.25E-01 & 1.95E-01 & 4.13E-01 & 3.91E+00 & 9.68E-01 & 2.62E-01 \\
										   & $T_\max$ (s) & \multicolumn{4}{|c|}{6.96E-01}						& \multicolumn{4}{|c|}{3.67E-01}						\\
	\hline
	\multirow{3}{*}{10}  & $N_p$        & 5.56E+04 & 1.63E+05 & 7.57E+04 & 6.07E+03 & 6.65E+04 & 1.33E+05 & 7.57E+04 & 4.50E+03 \\
										   & $N_m$ 			  & 17 			 & 387 		  & 46.73 	 & 31.78 		& 19 			 & 393 			& 46.68 	 & 27.93 		\\
										   & $V_m$ (MB)   & 4.01E-01 & 3.06E+00 & 8.40E-01 & 2.21E-01 & 4.40E-01 & 4.80E+00 & 2.14E+00 & 9.88E-01 \\
											 & $T_\max$ (s) & \multicolumn{4}{|c|}{7.55E-01}						& \multicolumn{4}{|c|}{3.30E-01}						\\
	\hline
	\multirow{3}{*}{100} & $N_p$        & 8.28E+02 & 4.98E+05 & 7.57E+04 & 3.14E+04 & 6.30E+04 & 1.10E+05 & 7.57E+04 & 4.20E+03 \\
										   & $N_m$ 			  & 11 			 & 373 		  & 41.18 	 & 30.85 		& 16 			 & 357 			& 41.77 	 & 22.66 		\\
										   & $V_m$ (MB)   & 5.28E-01 & 1.01E+01 & 2.65E+00 & 9.24E-01 & 9.76E-01 & 4.86E+00 & 3.78E+00 & 5.69E-01 \\
											 & $T_\max$ (s) & \multicolumn{4}{|c|}{9.25E-01}						& \multicolumn{4}{|c|}{3.55E-01}						\\
	\hline
	\end{tabular}
	}
	\caption{Detailed load balancing and communication information for the advection test problem for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$. Note how increasing the CFL number causes load imbalance and increases the communication volume while it does not affect the number of sent and received messages during an iteration of Semi-Lagrangian step.}
	\label{tab:communication}
\end{table}

\subsection{Reinitialization} \label{section::scaling_reinitialization}

Finally we present scaling results of our parallel reinitialization algorithm where we extensively make use of algorithm \ref{alg:overlap} for overlapping computation with communication when computing spatial derivatives. Our test is consisted of computing the signed distance to a collection of 100 sphere, whose radii and centers have been selected randomly (c.f. \ref{fig:reinit}). Computation is conducted on a forest with a $3\times3\times3$ macromesh and level-10 octree per macro-cell which results in a final grid with roughly $350$M nodes.
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/reinit.png}
\label{fig:reinit}
\caption{Computational domain for the reinitialization test}
\end{figure}
